Precise predictions for same-sign W-boson scattering at the LHC

### 2018-06-15 Draft_EPJC/article.tex
fixces related to Pascal's comment

The VBS process involving two same-sign $\PW$ bosons has the largest signal-to-background ratio of all the VBS processes at the Large Hadron Collider (LHC):
evidence for it was found at the centre-of-mass energy of $8~\TeV$ already~\cite{Aad:2014zda,Khachatryan:2014sta},
and it has been recently observed~\cite{Sirunyan:2017ret} and measured~\cite{Aaboud:2016ffv} 
at $13\TeV$ as well.
Presently, the measurements of VBS processes are limited by statistics, but the situation will change in the near future.
On the theoretical side, 
it is thus of prime importance to provide predictions with systematic uncertainties
at least comparable to the current and envisaged experimental precision~\cite{CMS:2016rcn,ATL-PHYS-PUB-2017-023}.

@@

      \item WHIZARD~\cite{Kilian:2007gr,Moretti:2001zz} is a multi-purpose event generator with LO matrix-element generator O'Mega. 
      For QCD amplitudes it uses the colour flow formalism~\cite{Kilian:2012pz}.
      For NLO QCD calculations~\cite{Nejad:2016bci}, where WHIZARD is in the final validation phase, it provides FKS subtraction terms~\cite{Frixione:1995ms,Frixione:1997np}, while virtual matrix elements are provided externally by OpenLoops~\cite{Cascioli:2011va} or Recola~\cite{Actis:2012qn,Actis:2016mpe}.
      Furthermore, WHIZARD can automatically provide POWHEG matching to parton shower~\cite{Reuter:2016qbi}.
      WHIZARD allows to simulate a huge number of BSM models as well, in particular in terms of higher-dimensional operators for VBS processes including means to provide unitarity limits~\cite{Alboteanu:2008my,Kilian:2014zja}.

@@

    \begin{table*}
        \footnotesize
        \begin{tabularx}{\textwidth}{c|c|X|X|X|X|X}
            Code  &  $\mathcal O(\alpha^6)$ $s, t, u$  &  $\mathcal O(\alpha^6)$ interf.  &  Non-res.  & NLO &  NF QCD  &  EW corr. to order $\mathcal O(\alphas \alpha^5)$  \\
            \hline
            \hline
            {\sc PHANTOM}       &  $s,t,u$  &  Yes      &  Yes              &  No    & -        & - \\
            {\sc Whizard}       &  $s,t,u$  &  Yes      &  Yes              &  No    & -        & - \\
            {\sc Bonsay}        &  $t,u$    &  No       &  Yes, virt. No    &  Yes   & No       &  No  \\
            {\sc Powheg}        &  $t,u$    &  No       &  Yes              &  Yes   & No       &  No  \\
            {\sc VBFNLO}        &  $s,t,u$  &  No       &  Yes              &  Yes   & No       &  No  \\
            {\sc MG5\_aMC}      &  $s,t,u$  &  Yes      &  Yes              &  Yes   & virt. No &  No \\
            {\sc MoCaNLO+Recola}&  $s,t,u$  &  Yes      &  Yes              &  Yes   & Yes      &  Yes  \\
        \end{tabularx}
        \caption{\label{tab:wg1_codes} Summary of the different properties of the computer programs employed in the comparison.}
    \end{table*}
    %
    \subsection{Input parameters}
        \label{subsec:inputpar}
    The hadronic scattering processes are simulated at the LHC with a centre-of-mass energy $\sqrt s = 13 \TeV$.
    The NNPDF~3.0 parton distribution functions~(PDFs)~\cite{Ball:2014uwa} with five massless flavours,\footnote{For the process considered, no bottom (anti-)quarks appear in the initial or final state at LO and NLO, as they would lead to top quarks, and not light jets, in the final state.} 
    NLO-QCD evolution, and a strong coupling constant $\alphas\left( \MZ \right) = 0.118$\footnote{Note that the {\sc Powheg-Box} uses its own implementation of the two loop running for $\alpha_{\rm s}$.} are employed.\footnote{The corresponding identifier {\tt lhaid} in the program LHAPDF6~\cite{Buckley:2014ana} is 260000.}
    Initial-state collinear singularities are factorised according to the ${\overline{\rm MS}}$ scheme, consistently with what is done in NNPDF.

@@

    \subsection{Comparison in the fiducial region}
    In Tab.~\ref{tab:wg1_LOrates}, we report the total rates at LO accuracy at order $\mathcal O (\alpha^6)$ obtained in the fiducial region 
    described in Sec.~\ref{subsec:inputpar}. Two things should be highlighted here: first, despite the different underlying approximations, 
    the two most-distant predictions ({\sc Powheg-box}\ and {\sc MG5\_aMC}) are only $0.7\%$ apart. This simply means that the details 
    of the various VBS approximations have an impact below $1\%$ at the level of the fiducial cross section at LO for a 
    typical phase-space volume used by experimental collaborations. This is in agreement with the findings of 
    Refs.~\cite{Denner:2012dz,Oleari:2003tc}. Second, the four complete predictions ({\sc Whizard}, {\sc PHANTOM}, {\sc MG5\_aMC}, 
    and {\sc MoCaNLO+Recola}) are not in statistical agreement. While we have checked the point-wise agreement of the matrix-element, 
    we cannot exclude other reasons for the disagreement, for example a non representative (\emph{i.e.}\ too-aggressive) 
    estimate of the Monte Carlo uncertainty or a non-perfect mapping of the six-body phase-space. However, the level of ambiguity is at the $0.5\%$ level, which 
    we deem satisfactory compared to the larger differences observed at NLO or when including matching to parton shower. 
    \begin{table}
        \centering
        \begin{tabular}{c|r@{ $\pm$ }l}
          Code  &  \multicolumn{2}{c}{$\sigma[\rm{fb}]$}  \\
            \hline
            \hline
            {\sc Bonsay}  &  $1.43636$ & $0.00002$ \\
            {\sc Powheg-Box}  &  $1.44092$ & $0.00009$ \\
            {\sc VBFNLO}  &  $1.43796$ & $0.00005$ \\
            {\sc PHANTOM} &  $1.4374\phantom{0}$ & $0.0006 $  \\
            {\sc Whizard} &  $1.4381\phantom{0}$ & $0.0002 $ \\
            {\sc MG5\_aMC}&  $1.4304\phantom{0}$ & $0.0007$ \\
            {\sc MoCaNLO+Recola}  &  $1.43476$ & $0.00009$ 
        \end{tabular}
        \caption{\label{tab:wg1_LOrates} Cross sections at LO accuracy and order $\mathcal{O}(\alpha^6)$.
        The complete $2\to 6$ matrix-element, without any approximation, is employed by {\sc PHANTOM},
        {\sc Whizard}, {\sc MG5\_aMC} and {\sc MoCaNLO+Recola}. The predictions are obtained 
        in the fiducial region described in Sec.~\protect\ref{subsec:inputpar}.
        The uncertainties shown refer to the estimated statistical errors of the Monte Carlo integrations.}
    \end{table}

@@

\section{Matching to parton shower}
    \label{sec:matching}
We now discuss how different predictions compare when the matching to parton shower is included, both at LO 
(\emph{i.e.}\ at order $\mathcal O (\alpha^6)$) and at NLO-QCD (\emph{i.e.}\ at order $\mathcal O (\alpha^6\alphas)$) accuracy. For such
a comparison we expect larger discrepancies than what we found at fixed order, as a consequence of the different
matching schemes, parton showers employed, and of other details of the matching (such as the choice of the parton shower initial scale). Among
the codes capable of providing fixed-order results, presented before, {\sc MG5\_aMC}, the {\sc Powheg-Box}, and {\sc VBFNLO}
can also provide results at (N)LO+PS accuracy. For {\sc VBFNLO} matched to {\sc Herwig} and the {\sc Powheg-Box}, we
restrict ourselves to show results only in the VBS approximation,
{\emph i.e.}\ the $s$-channel contributions are neglected here. Besides,
also {\sc PHANTOM} and {\sc WHIZARD} are used for LO+PS results.\\
{\sc MG5\_aMC}, which employs the {\sc MC@NLO}~\cite{Frixione:2002ik} matching procedure, is used together with {\sc Pythia8}~\cite{Sjostrand:2014zea} (version 8.223)
and {\sc Herwig7}~\cite{Bellm:2015jjp,Bellm:2017bvx} (version 7.1.2). For the latter, the default angular-ordered shower is employed. 
The same parton showers are employed for the LO results of {\sc PHANTOM}.
{\sc Pythia8} is also employed for the LO results of {\sc WHIZARD}. For the {\sc Powheg-Box}, the namesake
matching procedure is employed~\cite{Nason:2004rx,Frixione:2007vw}, together with {\sc Pythia8} (version 8.230). {\sc VBFNLO} serves as a matrix-element and phase-space provider
for the {\sc Matchbox} module~\cite{Platzer:2011bc} of {\sc
Herwig7}~\cite{Bellm:2015jjp,Bellm:2017bvx}, using an extended version of the Binoth Les Houches Accord
interface~\cite{Binoth:2010xt,Alioli:2013nda,Andersen:2014efa}. The {\sc Matchbox} module makes it
possible to choose between {\sc MC\-@NLO}-like and {\sc Powheg}-like
matching. As parton show\-er, both the default angular-ordered shower as
well as the dipole shower can be employed. We use here the subtractive (MC@NLO-type) matching to these parton-show\-er algorithms.
Whenever {\sc Pythia8}\ is used, the Monash tune~\cite{Skands:2014pea} is selected. Multiple-parton interactions are disabled.

@@

\begin{table}
    \centering
    \begin{tabular}{c|l@{ $\pm$ }l}
      Code  &  \multicolumn{2}{c}{$\sigma[\rm{fb}]$}  \\
        \hline\hline
        {\sc MG5\_aMC}+{\sc Pythia8}&  $1.352 $ & $0.003$  \\
        {\sc MG5\_aMC}+{\sc Herwig7}&  $1.342 $ & $ 0.003$  \\
        {\sc MG5\_aMC}+{\sc Pythia8}, $\Gamma_{\rm resc}$&  $1.275$ & $0.003$  \\
        {\sc MG5\_aMC}+{\sc Herwig7}, $\Gamma_{\rm resc}$&  $1.266$ & $ 0.003$  \\
        {\sc PHANTOM}+{\sc Pythia8} &  $1.235 $ & $0.001$  \\
        {\sc PHANTOM}+{\sc Herwig7} &  $1.258 $ & $0.001$  \\
        {\sc VBFNLO}+{\sc Herwig7-Dipole} &  $1.3001$ & $0.0002$  \\
        {\sc WHIZARD}+{\sc Pythia8} &  $1.229$ & $0.001$  \\
    \end{tabular}
    \caption{\label{tab:PSratesLO} Cross sections at LO+PS accuracy.
    The {\sc MG5\_aMC} results with $\Gamma_{\rm resc}$
    are rescaled to account for the effect related to the W-boson width computed by {\sc MadSpin} (see the text for details).
    The uncertainties shown refer to estimated statistical errors of the Monte Carlo integrations.}
\end{table}

>>>>

The VBS process involving two same-sign $\PW$ bosons has the largest signal-to-background ratio of all the VBS processes at the Large Hadron Collider (LHC):
evidence for it was found at the centre-of-mass energy of $8~\TeV$ already~\cite{Aad:2014zda,Khachatryan:2014sta},
and it has been recently measured~\cite{Aaboud:2016ffv} and (re-)observed~\cite{Sirunyan:2017ret}
at $13\TeV$ as well.
Presently, the measurements of VBS processes are limited by statistics, but the situation will change in the near future.
On the theoretical side, 
it is thus of prime importance to provide predictions with systematic uncertainties
at least comparable to the current and envisaged experimental precision~\cite{CMS:2016rcn,ATL-PHYS-PUB-2017-023}.

@@

      \item Whizard~\cite{Kilian:2007gr,Moretti:2001zz} is a multi-purpose event generator with LO matrix-element generator O'Mega. 
      For QCD amplitudes it uses the colour flow formalism~\cite{Kilian:2012pz}.
      For NLO QCD calculations~\cite{Nejad:2016bci}, where Whizard is in the final validation phase, it provides FKS subtraction terms~\cite{Frixione:1995ms,Frixione:1997np}, while virtual matrix elements are provided externally by OpenLoops~\cite{Cascioli:2011va} or Recola~\cite{Actis:2012qn,Actis:2016mpe}.
      Furthermore, Whizard can automatically provide POWHEG matching to parton shower~\cite{Reuter:2016qbi}.
      Whizard allows to simulate a huge number of BSM models as well, in particular in terms of higher-dimensional operators for VBS processes including means to provide unitarity limits~\cite{Alboteanu:2008my,Kilian:2014zja}.

@@

    \begin{table*}
        \footnotesize
        \begin{tabularx}{\textwidth}{c|c|X|X|X|X|X}
            Code  &  $\mathcal O(\alpha^6)$ $s, t, u$  &  $\mathcal O(\alpha^6)$ interf.  &  Non-res.  & NLO &  NF QCD  &  EW corr. to order $\mathcal O(\alphas \alpha^5)$  \\
            \hline
            \hline
            {\sc Phantom}       &  $s,t,u$  &  Yes      &  Yes              &  No    & -        & - \\
            {\sc Whizard}       &  $s,t,u$  &  Yes      &  Yes              &  No    & -        & - \\
            {\sc Bonsay}        &  $t,u$    &  No       &  Yes, virt. No    &  Yes   & No       &  No  \\
            {\sc Powheg}        &  $t,u$    &  No       &  Yes              &  Yes   & No       &  No  \\
            {\sc VBFNLO}        &  $s,t,u$  &  No       &  Yes              &  Yes   & No       &  No  \\
            {\sc MG5\_aMC}      &  $s,t,u$  &  Yes      &  Yes              &  Yes   & virt. No &  No \\
            {\sc MoCaNLO+Recola}&  $s,t,u$  &  Yes      &  Yes              &  Yes   & Yes      &  Yes  \\
        \end{tabularx}
        \caption{\label{tab:wg1_codes} Summary of the different properties of the computer programs employed in the comparison.}
    \end{table*}
    %
    \subsection{Input parameters}
        \label{subsec:inputpar}
    The hadronic scattering processes are simulated at the LHC with a centre-of-mass energy $\sqrt s = 13 \TeV$.
    The NNPDF~3.0 parton distribution functions~(PDFs)~\cite{Ball:2014uwa} with five massless flavours,\footnote{For the process considered, no bottom (anti-)quarks appear in the initial or final state at LO and NLO, as they would lead to top quarks, and not light jets, in the final state.} 
    NLO-QCD evolution, and a strong coupling constant $\alphas\left( \MZ \right) = 0.118$\footnote{Note that the {\sc Powheg-Box} uses its own implementation of the two loop running for $\alpha_{\rm s}$.} are employed.\footnote{The corresponding identifier {\tt lhaid} in the program LHAPDF6~\cite{Buckley:2014ana} is 260000.}
    Initial-state collinear singularities are factorised according to the ${\overline{\rm MS}}$ scheme, consistently with what is done in NNPDF.

@@

    \subsection{Comparison in the fiducial region}
    In Tab.~\ref{tab:wg1_LOrates}, we report the total rates at LO accuracy at order $\mathcal O (\alpha^6)$ obtained in the fiducial region 
    described in Sec.~\ref{subsec:inputpar}. Two things should be highlighted here: first, despite the different underlying approximations, 
    the two most-distant predictions ({\sc Powheg-box}\ and {\sc MG5\_aMC}) are only $0.7\%$ apart. This simply means that the details 
    of the various VBS approximations have an impact below $1\%$ at the level of the fiducial cross section at LO for a 
    typical phase-space volume used by experimental collaborations. This is in agreement with the findings of 
    Refs.~\cite{Denner:2012dz,Oleari:2003tc}. Second, the four complete predictions ({\sc Whizard}, {\sc Phantom}, {\sc MG5\_aMC}, 
    and {\sc MoCaNLO+Recola}) are not in statistical agreement. While we have checked the point-wise agreement of the matrix-element, 
    we cannot exclude other reasons for the disagreement, for example a non representative (\emph{i.e.}\ too-aggressive) 
    estimate of the Monte Carlo uncertainty or a non-perfect mapping of the six-body phase-space. However, the level of ambiguity is at the $0.5\%$ level, which 
    we deem satisfactory compared to the larger differences observed at NLO or when including matching to parton shower. 
    \begin{table}
        \centering
        \begin{tabular}{c|r@{ $\pm$ }l}
          Code  &  \multicolumn{2}{c}{$\sigma[\rm{fb}]$}  \\
            \hline
            \hline
            {\sc Bonsay}  &  $1.43636$ & $0.00002$ \\
            {\sc Powheg-Box}  &  $1.44092$ & $0.00009$ \\
            {\sc VBFNLO}  &  $1.43796$ & $0.00005$ \\
            {\sc Phantom} &  $1.4374\phantom{0}$ & $0.0006 $  \\
            {\sc Whizard} &  $1.4381\phantom{0}$ & $0.0002 $ \\
            {\sc MG5\_aMC}&  $1.4304\phantom{0}$ & $0.0007$ \\
            {\sc MoCaNLO+Recola}  &  $1.43476$ & $0.00009$ 
        \end{tabular}
        \caption{\label{tab:wg1_LOrates} Cross sections at LO accuracy and order $\mathcal{O}(\alpha^6)$.
        The complete $2\to 6$ matrix-element, without any approximation, is employed by {\sc Phantom},
        {\sc Whizard}, {\sc MG5\_aMC} and {\sc MoCaNLO+Recola}. The predictions are obtained 
        in the fiducial region described in Sec.~\protect\ref{subsec:inputpar}.
        The uncertainties shown refer to the estimated statistical errors of the Monte Carlo integrations.}
    \end{table}

@@

\section{Matching to parton shower}
    \label{sec:matching}
We now discuss how different predictions compare when the matching to parton shower is included, both at LO 
(\emph{i.e.}\ at order $\mathcal O (\alpha^6)$) and at NLO-QCD (\emph{i.e.}\ at order $\mathcal O (\alpha^6\alphas)$) accuracy. For such
a comparison we expect larger discrepancies than what we found at fixed order, as a consequence of the different
matching schemes, parton showers employed, and of other details of the matching (such as the choice of the parton shower initial scale). Among
the codes capable of providing fixed-order results, presented before, {\sc MG5\_aMC}, the {\sc Powheg-Box}, and {\sc VBFNLO}
can also provide results at (N)LO+PS accuracy. For {\sc VBFNLO} matched to {\sc Herwig} and the {\sc Powheg-Box}, we
restrict ourselves to show results only in the VBS approximation,
{\emph i.e.}\ the $s$-channel contributions are neglected here. Besides,
also {\sc Phantom} and {\sc Whizard} are used for LO+PS results.\\
{\sc MG5\_aMC}, which employs the {\sc MC@NLO}~\cite{Frixione:2002ik} matching procedure, is used together with {\sc Pythia8}~\cite{Sjostrand:2014zea} (version 8.223)
and {\sc Herwig7}~\cite{Bellm:2015jjp,Bellm:2017bvx} (version 7.1.2). For the latter, the default angular-ordered shower is employed. 
The same parton showers are employed for the LO results of {\sc Phantom}.
{\sc Pythia8} is also employed for the LO results of {\sc Whizard}. For the {\sc Powheg-Box}, the namesake
matching procedure is employed~\cite{Nason:2004rx,Frixione:2007vw}, together with {\sc Pythia8} (version 8.230). {\sc VBFNLO} serves as a matrix-element and phase-space provider
for the {\sc Matchbox} module~\cite{Platzer:2011bc} of {\sc
Herwig7}~\cite{Bellm:2015jjp,Bellm:2017bvx}, using an extended version of the Binoth Les Houches Accord
interface~\cite{Binoth:2010xt,Alioli:2013nda,Andersen:2014efa}. The {\sc Matchbox} module makes it
possible to choose between {\sc MC\-@NLO}-like and {\sc Powheg}-like
matching. As parton show\-er, both the default angular-ordered shower as
well as the dipole shower can be employed. We use here the subtractive (MC@NLO-type) matching to these parton-show\-er algorithms.
Whenever {\sc Pythia8}\ is used, the Monash tune~\cite{Skands:2014pea} is selected. Multiple-parton interactions are disabled.

@@

\begin{table}
    \centering
    \begin{tabular}{c|l@{ $\pm$ }l}
      Code  &  \multicolumn{2}{c}{$\sigma[\rm{fb}]$}  \\
        \hline\hline
        {\sc MG5\_aMC}+{\sc Pythia8}&  $1.352 $ & $0.003$  \\
        {\sc MG5\_aMC}+{\sc Herwig7}&  $1.342 $ & $ 0.003$  \\
        {\sc MG5\_aMC}+{\sc Pythia8}, $\Gamma_{\rm resc}$&  $1.275$ & $0.003$  \\
        {\sc MG5\_aMC}+{\sc Herwig7}, $\Gamma_{\rm resc}$&  $1.266$ & $ 0.003$  \\
        {\sc Phantom}+{\sc Pythia8} &  $1.235 $ & $0.001$  \\
        {\sc Phantom}+{\sc Herwig7} &  $1.258 $ & $0.001$  \\
        {\sc VBFNLO}+{\sc Herwig7-Dipole} &  $1.3001$ & $0.0002$  \\
        {\sc Whizard}+{\sc Pythia8} &  $1.229$ & $0.001$  \\
    \end{tabular}
    \caption{\label{tab:PSratesLO} Cross sections at LO+PS accuracy.
    The {\sc MG5\_aMC} results with $\Gamma_{\rm resc}$
    are rescaled to account for the effect related to the W-boson width computed by {\sc MadSpin} (see the text for details).
    The uncertainties shown refer to estimated statistical errors of the Monte Carlo integrations.}
\end{table}
